Notes on conformal manifolds

### 2025-05-26 Draft_MLFlatDirections.tex
update the draft which now looks almost like a v0 of the paper. Removed the itemize, add an intro and a conclusion. Write a sugra setup section

@@

\section{Supergravity setup}
\begin{itemize}[label=\textbullet]
	\item Context: 6d ${\rm AdS}_{3}\times S^{3}$, truncation to 3d, potential, conformal manifold.
	\item Choice of truncation from 32 to 13 to 5 variables.
\end{itemize}

\paragraph{Potential}
The 22 scalars of the theory can be parametrised by ($22 = 32 - 10$, with 10 scalars gauge fixed using translations in the gauge group)
\begin{itemize}[label=\textbullet]
	\item $m = \nu\nu^{T}\in{\rm GL}(3,\mathbb{R})$ parametrizing the coset ${\rm GL}(3,\mathbb{R})/{\rm SO}(3)$,
	\item $\phi$ a $3\times3$ antisymmetric matrix,
	\item $\xi$ a $3\times4$ matrix, and $\xi^{2} = \xi \xi^{T}$,
	\item a dilaton $\tilde{\varphi}$.
\end{itemize}
We can further restrict ourselves to a set of 13 scalars (see ref.~\cite{Eloy:2021fhc}), with
\begin{equation}
	\begin{aligned}
		\xi &= \begin{pmatrix}
					0 & 0 & 0 & x_{1} \\
					0 & 0 & 0 & x_{2} \\
					0 & 0 & 0 & x_{3}
				\end{pmatrix},
		\quad
		\phi = \begin{pmatrix}
					0 & x_{4} & x_{5} \\
					-x_{4} & 0 & x_{6} \\
					-x_{5} & -x_{6} & 0
				\end{pmatrix},
		\quad
		\tilde{\varphi} = x_{13} \\[5pt]
		\nu &= e^{(6\,x_{7}+3\,\tilde{x}_{8}+\sqrt{3}\,x_{9})/6}
				\begin{pmatrix}
					1 & \frac{x_{10}}{\sqrt{2}} & \frac{x_{11}}{\sqrt{2}} + \frac{x_{10}x_{12}}{4} \\
					0 & e^{-\tilde{x}_{8}} & \frac{e^{-\tilde{x}_{8}}\,x_{12}}{\sqrt{2}} \\
					0 & 0 & e^{-(\tilde{x}_{8}+\sqrt{3}\,x_{9})/2}
				\end{pmatrix}.
	\end{aligned}
\end{equation}
Scalar potential from ref.~\cite{Eloy:2021fhc}:
%
\begin{equation} \label{eq:scalarpotential}
	\begin{aligned}
 		V & = 4\,e^{-4\tilde\varphi}+2\,e^{-2\tilde\varphi}\Big[-\tr\left(m+m^{-1}\right)+\tr\left(\phi m^{-1}\phi\right) -2\,\tr\left(\phi m^{-1}\xi^{2}\right)-2\,\tr\left(\xi^{2}\right)\\
 		&\qquad\quad-\tr\left(\xi^{2}m^{-1}\xi^{2}\right)  +\frac{1}{2}\,\det\left(m^{-1}\right)\left(1-\tr\left(\phi^{2}\right)-\tr\left(\xi^{4}\right)+\tr\left(\xi^{2}\right)^{2}\right) \\
 		&\qquad\quad +\frac{1}{2}\,{\rm T}\left(m^{-1}(\xi^{2}-\phi),(\xi^{2}+\phi)m^{-1},m+(\xi^{2}+\phi)m^{-1}(\xi^{2}-\phi)+2\,\xi^{2}\right)\\
 		&\qquad\quad +\frac{1}{4}\,{\rm T}\left(m^{-1},m+(\xi^{2}+\phi)m^{-1}(\xi^{2}-\phi)+2\,\xi^{2},m+(\xi^{2}+\phi)m^{-1}(\xi^{2}-\phi)+2\,\xi^{2}\right)\Big], 
 	\end{aligned}
\end{equation}
where ${\rm T}\left(A,B,C\right)=\varepsilon_{mnp}\,\varepsilon_{qrs}\,A^{mq}B^{nr}C^{ps}$.

As a first simplified example we have considered only the parameters $x_{1}, x_{2}, x_{4}, \tilde{x}_{8}$ and $x_{10}$. The potential~\eqref{eq:scalarpotential} becomes
\begin{equation} \label{eq:scalarpotential124810}
	\begin{aligned}
		V &= \frac{1}{8}\, e^{-2\,\tilde{x}_{8}} \bigg[4 + 4\,x_{1}^4 + e^{4\,\tilde{x}_{8}} \big(2 + x_{10}^2\big)^2 \big(1 + x_{2}^4\big)\\
		  & \quad\qquad\qquad - 4\,e^{3\,\tilde{x}_{8}}\,\big(2 + x_{10}^2\big) \Big(2 - x_{1}^2 + \sqrt{2}\,x_{1} x_{10} x_{2}^3 + x_{2}^4 + 
      			2\,x_{1} x_{2} x_{4} + x_{4}^2 - x_{2}^2 \big(1 - x_{1}^2 + x_{4}^2\big)\Big)\\
        & \quad\qquad\qquad - 8\,e^{\tilde{x}_{8}}\,\Big(2 + x_{1}^4 + \sqrt{2}\,x_{1}^3 x_{10} x_{2} - x_{2}^2 - 2\,x_{1} x_{2} x_{4} + 
      			x_{4}^2 - x_{1}^2 \big(1 - x_{2}^2 + x_{4}^2\big)\Big)\\
        & \quad\qquad\qquad + 4\,e^{2\,\tilde{x}_{8}}\,\Big(2 - 4\,x_{1}^2 + x_{1}^4 - 4\,x_{2}^2 + 4\,x_{1}^2 x_{2}^2 + x_{2}^4
        			+ x_{10}^2\big(1 + 3\,x_{1}^2 x_{2}^2\big) + 4\,x_{4}^2 + x_{4}^4\Big)\\
        & \quad\qquad\qquad + 8\sqrt{2}\,x_{10}\,e^{2\,\tilde{x}_{8}}\,\Big(x_{1}^3 x_{2} + x_{1}^2 x_{4} - x_{2}^2 x_{4} + x_{1} \big(x_{2}^3 - x_{2} x_{4}^2\big)\Big)\bigg].
	\end{aligned}
\end{equation}
\ce{Change notation to keep $x$ for the polynomial variables?}

\ce{Discuss here why we need numerical methods: to complicated for mathematica to simplify the gradient.}

\section{Numerical analysis}
	\subsection{Gradient descent and local analysis}
		\paragraph{Gradient descent: implementation, sampling, time needed, loss}

@@

		\begin{figure}
			\centering
			\includegraphics[scale=0.35]{Figures/nabla_V_thres.png}
			\quad
			\includegraphics[scale=0.35]{Figures/V_values_thres.png}
			\caption{Left panel: Number of data points with gradient norm smaller than the threshold indicated on the x-axis. Right panel: Number of data points whose value deviates from $-4$ by less than the threshold shown on the x-axis.}
			\label{fig:Analyze_V_points}
		\end{figure}

@@

		\begin{figure}
			\centering
			\includegraphics[scale = 0.5]{Figures/Loss_grad_des.png}
			\caption{Evolution of the loss function during gradient descent and corresponding learning rate schedule, both plotted on a logarithmic scale. Red dashed lines indicate epochs at which the optimizer was reinitialized.}
			\label{fig:loss_grad_des}
		\end{figure}

@@

		We can make a few comments on those plots. First, we observe non-trivial correlations in the data. Some of these may result from larger basins of attraction, such as in the $x_1/x_2$ plot. However, the structures seen in the $x_4/\tilde{x}_8$ or $x_4/x_{10}$ plots likely reflect genuine features of the manifold, and will show latter that this is indeed the case. We also note that alll directions appear to be well populated. 
		%Furthermore, we can notice that a large majority of the points converge to the origin $\vec{0}$. This is not surprising, as this just mean that the round sphere solution has a larger basin of attraction, which sounds reasonable. We do not however attempt to shift the points along flat directions to depopulate the origin or repopulate other regions.

		\paragraph{Local PCA: graphs of number of patches with given local dimension, and with varying size of local patch, discuss optimal patch size (we want at least few points in each direction)}

		Once the gradient descent has been completed and the flat direction sampled, the next step is to identify the structure of the underlying manifold. Our goal is to eventually obtain an analytical expression, not just a numerical description. Before applying symbolic regression to search for such an expression, we can first perform some exploratory analyses to better understand the data. Specifically, we aim to determine the dimensionality of the manifold and whether it consists of a single connected component or multiple disjoint components (e.g., two intersecting hyperplanes).

@@

		\begin{figure}
			\centering
			\includegraphics[scale=0.5]{Figures/Local_PCA.png}
			\caption{Results of the local PCA analysis. The $x$-axis shows the dimension inferred by the algorithm, and the $y$-axis indicates the proportion of points for which that dimension was found. Each curve corresponds to a different value of $k$ in the $k$-nearest neighbours.}
			\label{local_pca}
		\end{figure}

		We observe that for every choice of $k$, there is a prominent peak at $d = 3$, suggesting that the underlying manifold is three-dimensional. For $k = 5$, a noticeable fraction of points are assigned a dimensionality of 2. This can be attributed to the fact that if the true dimension is 3, then selecting only 5 neighbours may not sufficiently populate all three directions, leading the algorithm to underestimate the dimensionality.

@@

		\paragraph{Clustering: HDBScan (only parameter: minimal number of points in a cluster), and graphs of 3d slices of the 5d space}

@@

		\begin{figure}
			\centering
			\includegraphics[scale=0.4]{Figures/3d_plot_x4_x8_x10.png}
			\includegraphics[scale=0.4]{Figures/3d_plot_x2_x4_x8.png}
			\caption{3D plots of the data in selected coordinates. Left: $(x_4, x_8, x_{10})$. Right: $(x_2, x_4, x_8)$.}
			\label{3dplots}
		\end{figure}

@@

	\subsection{Annealed Importance Sampling for polynomial symbolic regression}
	\begin{itemize}[label=\textbullet]
		\item We can convert the potential to a polynomial by converting the variables that appear in exponentials to logs.
		\item As the potential is a polynomial, the solutions satisfy polynomials equations. Discuss that fact that we could get polynomials directly by taking the gradient of the potential, but those will be to complicated to express the vacuum in a usable way.  We search those using Annealed Importance Sampling.

>>>>

\def\bbar#1{\bar{\bar{#1}}}

\newcommand{\cN}{{\cal N}}
\newcommand{\cA}{{\cal A}}
%\newcommand{\Tr}{\mbox{Tr}}
%\newcommand{\tr}{\mbox{tr}}
\newcommand{\cZ}{{\cal Z}}

\newcommand{\bK}{{\bar{K}}}
\newcommand{\bL}{{\bar{L}}}
\newcommand{\bM}{{\bar{M}}}
\newcommand{\bN}{{\bar{N}}}
\newcommand{\bP}{{\bar{P}}}
\newcommand{\bQ}{{\bar{Q}}}
\newcommand{\bR}{{\bar{R}}}
\newcommand{\bS}{{\bar{S}}}
\newcommand{\bk}{{\bar{k}}}
\newcommand{\bl}{{\bar{l}}}
\newcommand{\bzero}{{\bar{0}}}
\newcommand{\balpha}{{\bar{\alpha}}}
\newcommand{\bbeta}{{\bar{\beta}}}

\def\dA{{\dot{A}}}
\def\dB{{\dot{B}}}
\def\dC{{\dot{C}}}
\def\dD{{\dot{D}}}

\def\ft#1#2{{\textstyle{{\scriptstyle #1}\over {\scriptstyle #2}}}}
\def\fft#1#2{{#1 \over #2}}
\def\del{\partial}
\def\vp{\varphi}
\def\sst#1{{\scriptscriptstyle #1}}
\def\oneone{\rlap 1\mkern4mu{\rm l}}
\def\td{\tilde}
\def\wtd{\widetilde}

@@

The AdS/CFT correspondence stands as one of the most profound dualities in theoretical physics, establishing a remarkable equivalence between gravitational theories in Anti-de Sitter (AdS) spacetimes and conformal field theories (CFTs) living on their boundaries. This correspondence has revolutionized our understanding of both quantum gravity and strongly coupled field theories, providing unprecedented insights into the holographic nature of gravity.

Within this holographic framework, supergravity theories in AdS backgrounds serve as the low-energy effective descriptions of string theory compactifications, making them natural laboratories for exploring the gravitational side of the duality. On the CFT side, the dual conformal field theories often exhibit rich structures of exactly marginal operators whose coupling constants parametrize conformal manifolds—moduli spaces where the theory remains conformal but with varying coupling strengths. The AdS/CFT correspondence maps these conformal manifolds to flat directions in the supergravity scalar potential, where the cosmological constant remains fixed while scalar field configurations vary continuously.

Understanding these flat directions is therefore of great importance for several interconnected reasons. First, they encode the conformal manifolds of the dual CFTs, providing direct access to the space of exactly marginal deformations and their associated beta functions. Second, they reveal the vacuum structure of supergravity theories, determining which background configurations preserve supersymmetry and their associated stability properties.

However, the explicit characterization of these flat directions presents formidable technical challenges. Supergravity scalar potentials, even in truncated models, typically involve dozens of scalar fields with intricate non-linear interactions. The resulting expressions for critical points—where all first derivatives vanish—quickly become too complex for traditional symbolic manipulation, rendering analytical approaches computationally intractable. This complexity paradox is particularly frustrating given that physically meaningful solutions often admit relatively simple parametric descriptions once discovered.

The emergence of machine learning techniques in theoretical physics offers new avenues for tackling such problems. Rather than attempting to solve the full symbolic system directly, we can leverage numerical methods to sample the solution space and then employ symbolic regression to extract analytical patterns from the resulting data. This hybrid approach potentially circumvents the computational bottlenecks that plague purely symbolic methods while maintaining the goal of obtaining exact analytical expressions.

In this work, we demonstrate the viability of this machine learning approach by applying it to a five-scalar subsector of a 13 scalar consistent truncation of three-dimensional chiral $\mathcal{N} = (2,0)$ supergravity arising from $AdS_3 \times S^3$ compactification from 6d supergravity, or from the compactification of type IIB supergravity on $AdS_3 \times S^3 \times T^4$. Our methodology combines gradient descent sampling of the flat direction manifold with a novel annealed importance sampling algorithm for polynomial symbolic regression. We show that this approach successfully identifies the underlying three-dimensional conformal manifold and automatically discovers the polynomial constraints that characterize it analytically.

The paper is organized as follows. Section 2 establishes the supergravity setup, presenting the scalar potential in its full complexity and motivating the truncation to five fields. Section 3 details our numerical methodology, covering both the gradient descent sampling procedure and the annealed importance sampling approach to symbolic regression. Section 4 analyzes the discovered supergravity solutions, including their geometric interpretation and spectrum analysis. Finally, Section 5 concludes with prospects for extending this approach to higher-dimensional cases and its broader implications for systematic studies of conformal manifolds in holographic theories.

\section{Supergravity setup}
%\begin{itemize}[label=\textbullet]
%	\item Context: 6d ${\rm AdS}_{3}\times S^{3}$, truncation to 3d, potential, conformal manifold.

Three-dimensional $\mathcal{N}=8$ (half-maximal) gauged supergravity is governed by the Lagrangian
%
\begin{equation}	\label{eq: lagrangian_rephrased} 
	e^{-1}\mathcal{L}=R+\frac1{8}g^{\mu\nu}D_\mu M^{\bM\bN}D_\nu M_{\bM\bN}+e^{-1}\mathcal{L}_{\text{CS}}-V\,,
\end{equation}
%
which comprises the Einstein-Hilbert term $R$, the kinetic term for scalar fields parametrized by the matrix $M_{\bM\bN}$, a Chern-Simons contribution $\mathcal{L}_{\text{CS}}$, and a scalar potential $V$.

The gauging structure is encoded in an embedding tensor that takes the general form
%
\begin{equation}	\label{eq: embtensor_rephrased}
	\Theta_{\bK\bL\vert\bM\bN}=\theta_{\bK\bL\bM\bN}+\frac12\Big(\eta_{\bM[\bK}\theta_{\bL]\bN}-\eta_{\bN[\bK}\theta_{\bL]\bM}\Big)+\theta\,\eta_{\bM[\bK}\eta_{\bL]\bN}\,.
\end{equation}
%
where we have a fully antisymmetric component $\theta_{\bK\bL\bM\bN}=\theta_{[\bK\bL\bM\bN]}$, a symmetric traceless part $\theta_{\bL\bK}=\theta_{(\bL\bK)}$, and a constant parameter $\theta$. The metric $\eta_{\bK\bL}$ is the SO(8,4)-invariant bilinear form used for index contractions.

The scalar degrees of freedom parametrize the coset space
%
\begin{equation}	\label{eq: scalarcoset_rephrased}
	\frac{\text{SO}(8,4)}{\text{SO}(8)\times\text{SO}(4)}\,,
\end{equation}
%
through the symmetric matrix $M_{\bK\bL}=\mathcal{V}_{\bK}{}^{\bM}\mathcal{V}_{\bL}{}^{\bN}\delta_{\bM\bN}$, where $\mathcal{V}_{\bK}{}^{\bM}$ represents the coset representative.

The gauge covariant derivatives are constructed using the embedding tensor according to
%
\begin{equation}
	D_\mu =\partial_\mu + A_\mu{}^{\bM\bN}\,\Theta_{\bM\bN\vert\bP\bQ}\, T^{\bP\bQ}\,,
\end{equation}
%
where $A_\mu{}^{\bM\bN}$ are the gauge fields and $T^{\bP\bQ}$ are the generators of the SO(8,4) algebra satisfying
%
\begin{equation} \label{eq:so84gen_rephrased}
	\big(T^{\bar M\bar N}\big){}_{\bar P}{}^{\bar Q} = 2\,\delta_{\bar P}{}^{[\bar M}\,\eta^{\bar N]\bar Q}\,.
\end{equation}
%
The covariant derivative acting on the scalar matrix becomes
%
\begin{equation}
	D_\mu M_{\bM\bN}=\partial_\mu M_{\bM\bN}+4\,A_\mu{}^{\bP\bQ}\,\Theta_{\bP\bQ\vert(\bM}{}^{\bK}\, M_{\bN)\bK}\,,
\end{equation}
%
ensuring gauge invariance of the scalar kinetic terms.

The Chern-Simons contribution is given by
%
\begin{equation}
	{\cal L}_{\rm CS}= -\varepsilon^{\,\mu\nu\rho}\,\Theta_{\bar M\bar N|\bar P\bar Q}\,A_{\mu}{}^{\bar M\bar N}\left(\partial_{\nu}\,A_{\rho}{}^{\bar P\bar Q}  + \frac{1}{3}\, \Theta_{\bar R\bar S|\bar U\bar V}\,f^{\bar P\bar Q,\bar R\bar S}{}_{\bar X\bar Y}\, A_{\nu}{}^{\bar U\bar V} A_{\rho}{}^{\bar X\bar Y} \right)\,,
\end{equation}
%
where $f^{\bar M\bar N,\bar P\bar Q}{}_{\bar K\bar L} = 4\,\delta_{[\bar K}{}^{[\bar M}\eta^{\bar N][\bar P}\delta_{\bar L]}{}^{\bar Q]}$ are the structure constants of the SO(8,4) Lie algebra, and $\varepsilon^{\mu\nu\rho}$ is the three-dimensional Levi-Civita tensor.

The scalar potential can be expressed in terms of the embedding tensor components as
%
{\setlength\arraycolsep{1.2pt}
	\begin{equation}	\label{eq: scalarpot_rephrased}
		\begin{aligned}
			V	&=	\frac1{12}\,\theta_{\bK\bL\bM\bN}\theta_{\bP\bQ\bR\bS}\Big(M^{\bK\bP}M^{\bL\bQ}M^{\bM\bR}M^{\bN\bS}-6\,M^{\bK\bP}M^{\bL\bQ}\eta^{\bM\bR}\eta^{\bN\bS}\\
			&\qquad\qquad\qquad\qquad\quad+8\,M^{\bK\bP}\eta^{\bL\bQ}\eta^{\bM\bR}\eta^{\bN\bS}-3\,\eta^{\bK\bP}\eta^{\bL\bQ}\eta^{\bM\bR}\eta^{\bN\bS}\Big)\\
			&\quad +\frac1{8}\,\theta_{\bK\bL}\theta_{\bP\bQ}\Big(2\,M^{\bK\bP}M^{\bL\bQ}-2\,\eta^{\bK\bP}\eta^{\bL\bQ}-M^{\bK\bL}M^{\bP\bQ}\Big)+4\,\theta\theta_{\bK\bL}M^{\bK\bL}-32\,\theta^2\,.
		\end{aligned}
	\end{equation}
}
In order to describe the $\mathcal{N} = (2,0)$ theory, it is useful to break the SO(8,4) group as

\begin{equation}	\label{eq: gl3gradingbar}
	\begin{aligned}
		\text{SO}(8,4)	&\longrightarrow	\enspace \text{GL}(3,\mathbb{R})\times\text{SO}(1,1)\times\text{SO}(4)_{\rm global}\,,	\\
		X^{\bM}		&\longrightarrow	\quad \{X^{\bar m},\; X_{\bar m},\; X^\bzero,\; X_\bzero,\; X^\balpha\}\,,
	\end{aligned}
\end{equation}
where $\bar m\in\llbracket1,3\rrbracket$ and $\balpha\in\llbracket9,12\rrbracket$ label the SL(3, $\mathbb{R}$) and SO(4)$_{\rm global}$ vector representations, respectively. In this basis, the SO(8,4)-invariant tensor reads
%
\begin{equation}	\label{eq: Paulieta}
	\eta_{\bM\bN}=
	\begin{pmatrix}
		0 & \delta_{\bar m}{}^{\bar n} & 0 & 0 &0\\
		\delta^{\bar m}{}_{\bar n} & 0 & 0 & 0 &0\\
		0 & 0 & 0 & 1 & 0\\
		0 & 0 & 1 & 0 & 0\\
		0 & 0 & 0 & 0 & -\delta_{\balpha\bbeta}
	\end{pmatrix}\,.
\end{equation}
The embedding tensor have the the following expression
\begin{equation}{\label{eq:SO84embeddings}}
		\theta_{\bar m \bar n \bar p\,\bzero}=-2\,\lambda\,\varepsilon_{\bar m \bar n \bar p}\,,		\qquad	\theta_{\bar m \bar n \bar p}{}^{\bzero}=\varepsilon_{\bar m \bar n \bar p}\,, \qquad	\theta_{\bar m \bar n}{}^{\bar p}{}_{\bzero}=\varepsilon_{\bar m \bar n \bar p}\,, 
\end{equation}
or equivalently

\begin{equation}
	X_{\bar m\bar n \bar p} = \varepsilon_{\bar m \bar n \bar p}\,,		\qquad
	X_{\bar m}{}^{\bar n \bar p} = \varepsilon_{\bar m \bar n \bar p}\,,	\qquad
	X^{\bar m}{}_{\bar n}{}^{\bar p} = \varepsilon_{\bar m \bar n \bar p}\,,	\qquad
	X^{\bar m\bar n}{}_{\bar p} = \varepsilon_{\bar m \bar n \bar p}\,,
\end{equation}
	%item Choice of truncation from 32 to 13 to 5 variables.
%\end{itemize}
Parametrizing the scalar matrix following \cite{Eloy:2021fhc}, we have for the scalar matrix 
\begin{equation}	\label{eq: scalarmatrix}
	M_{\bM\bN}
	=
	\begin{pmatrix}
		m+(\xi^2+\phi)m^{-1}(\xi^2-\phi)+2\xi^2	& 	(\xi^2+\phi)m^{-1}		&	0	&  0  &	-\sqrt2\,[1+(\xi^2+\phi)m^{-1}]\xi	\\
		m^{-1}(\xi^2-\phi)	& 	m^{-1}	&	0  	&  0  &	-\sqrt2\,m^{-1}\xi	\\
		0		&	0		&	e^{2\tilde\varphi}	&	0				&	0		\\
		0		&	0		&	0				&	e^{-2\tilde\varphi}	&	0		\\
		-\sqrt2\,\xi^T[1+m^{-1}(\xi^2-\phi)]	&	-\sqrt2\,\xi^Tm^{-1}	&	0	&	0	&	1+2\,\xi^Tm^{-1}\xi	\\
	\end{pmatrix}\,,
\end{equation}
which encodes 22 scalars of the theory ($22 = 32 - 10$, with 10 scalars gauge fixed using translations in the gauge group), which have been parametrized by 
$m = \nu\nu^{T}\in{\rm GL}(3,\mathbb{R})$ parametrizing the coset ${\rm GL}(3,\mathbb{R})/{\rm SO}(3)$, $\phi$ a $3\times3$ antisymmetric matrix,$\xi$ a $3\times4$ matrix, and $\xi^{2} = \xi \xi^{T}$, a dilaton $\tilde{\varphi}$. With this parametrization, the potential takes the from ref.~\cite{Eloy:2021fhc}:
\begin{equation} \label{eq:scalarpotential}
	\begin{aligned}
		V & = 4\,e^{-4\tilde\varphi}+2\,e^{-2\tilde\varphi}\Big[-\tr\left(m+m^{-1}\right)+\tr\left(\phi m^{-1}\phi\right) -2\,\tr\left(\phi m^{-1}\xi^{2}\right)-2\,\tr\left(\xi^{2}\right)\\
		&\qquad\quad-\tr\left(\xi^{2}m^{-1}\xi^{2}\right)  +\frac{1}{2}\,\det\left(m^{-1}\right)\left(1-\tr\left(\phi^{2}\right)-\tr\left(\xi^{4}\right)+\tr\left(\xi^{2}\right)^{2}\right) \\
		&\qquad\quad +\frac{1}{2}\,{\rm T}\left(m^{-1}(\xi^{2}-\phi),(\xi^{2}+\phi)m^{-1},m+(\xi^{2}+\phi)m^{-1}(\xi^{2}-\phi)+2\,\xi^{2}\right)\\
		&\qquad\quad +\frac{1}{4}\,{\rm T}\left(m^{-1},m+(\xi^{2}+\phi)m^{-1}(\xi^{2}-\phi)+2\,\xi^{2},m+(\xi^{2}+\phi)m^{-1}(\xi^{2}-\phi)+2\,\xi^{2}\right)\Big], 
	\end{aligned}
\end{equation}
where ${\rm T}\left(A,B,C\right)=\varepsilon_{mnp}\,\varepsilon_{qrs}\,A^{mq}B^{nr}C^{ps}$. We can further restrict ourselves to a set of 13 scalars which is still a consistent truncation, with 
\begin{equation}
	\begin{aligned}
		\xi &= \begin{pmatrix}
					0 & 0 & 0 & x_{1} \\
					0 & 0 & 0 & x_{2} \\
					0 & 0 & 0 & x_{3}
				\end{pmatrix}, \\[5pt]
		\phi &= \begin{pmatrix}
					0 & x_{4} & x_{5} \\
					-x_{4} & 0 & x_{6} \\
					-x_{5} & -x_{6} & 0
				\end{pmatrix}, \\[5pt]
		\tilde{\varphi} &= \tilde{x}_{13}, \\[5pt]
		\nu &= e^{(6\,\tilde{x}_{7}+3\,\tilde{x}_{8}+\sqrt{3}\,\tilde{x}_{9})/6}
				\begin{pmatrix}
					1 & \frac{x_{10}}{\sqrt{2}} & \frac{x_{11}}{\sqrt{2}} + \frac{x_{10}x_{12}}{4} \\
					0 & e^{-\tilde{x}_{8}} & \frac{e^{-\tilde{x}_{8}}\,x_{12}}{\sqrt{2}} \\
					0 & 0 & e^{-(\tilde{x}_{8}+\sqrt{3}\,\tilde{x}_{9})/2}
				\end{pmatrix}.
	\end{aligned}
\end{equation}

As an initial simplification, we restrict our analysis to the scalar fields $x_{1}, x_{2}, x_{4}, \tilde{x}_{8}$, and $x_{10}$. This choice is made to facilitate a more tractable starting point for the investigation, and corresponds to a randomly selected subset of the original set of thirteen scalar fields. It is important to note that this subset no longer defines a consistent truncation of the full theory. Nevertheless, despite the fact that the effective theory defined by these five scalars may be formally ill-defined—due to potential higher-order interactions that could source the remaining scalar fields—the solutions we obtain remain valid solutions of the complete theory. This is because we compute the critical points of the scalar potential by first evaluating $\nabla V = 0$, and only then setting the remaining scalar fields to zero. By adopting this procedure—taking the gradient of the potential prior to truncation—we ensure that the resulting configurations satisfy the equations of motion and are thus genuine solutions of the full theory. Under this truncation, the scalar potential~\eqref{eq:scalarpotential} reduces to:
\begin{equation} \label{eq:scalarpotential124810}
	\begin{aligned}
		V &= \frac{1}{8}\, e^{-2\,\tilde{x}_{8}} \bigg[4 + 4\,x_{1}^4 + e^{4\,\tilde{x}_{8}} \big(2 + x_{10}^2\big)^2 \big(1 + x_{2}^4\big)\\
		  & \quad\qquad\qquad - 4\,e^{3\,\tilde{x}_{8}}\,\big(2 + x_{10}^2\big) \Big(2 - x_{1}^2 + \sqrt{2}\,x_{1} x_{10} x_{2}^3 + x_{2}^4 + 
      			2\,x_{1} x_{2} x_{4} + x_{4}^2 - x_{2}^2 \big(1 - x_{1}^2 + x_{4}^2\big)\Big)\\
        & \quad\qquad\qquad - 8\,e^{\tilde{x}_{8}}\,\Big(2 + x_{1}^4 + \sqrt{2}\,x_{1}^3 x_{10} x_{2} - x_{2}^2 - 2\,x_{1} x_{2} x_{4} + 
      			x_{4}^2 - x_{1}^2 \big(1 - x_{2}^2 + x_{4}^2\big)\Big)\\
        & \quad\qquad\qquad + 4\,e^{2\,\tilde{x}_{8}}\,\Big(2 - 4\,x_{1}^2 + x_{1}^4 - 4\,x_{2}^2 + 4\,x_{1}^2 x_{2}^2 + x_{2}^4
        			+ x_{10}^2\big(1 + 3\,x_{1}^2 x_{2}^2\big) + 4\,x_{4}^2 + x_{4}^4\Big)\\
        & \quad\qquad\qquad + 8\sqrt{2}\,x_{10}\,e^{2\,\tilde{x}_{8}}\,\Big(x_{1}^3 x_{2} + x_{1}^2 x_{4} - x_{2}^2 x_{4} + x_{1} \big(x_{2}^3 - x_{2} x_{4}^2\big)\Big)\bigg].
	\end{aligned}
\end{equation}

Let us make a comment at this point. Looking at the potential given in Eq.~\ref{eq:scalarpotential124810} and since we have an explicit formula, one might argue that the problem of identifying flat directions could be addressed by directly computing the gradient of the expression and subsequently simplifying the resulting equations. However, attempting to carry out this procedure using a symbolic solver such as Mathematica quickly reveals its limitations: the resulting expressions are too complex to be simplified into a manageable form, and do not yield usable relationships that express some variables in terms of others. This complexity does not, however, rule out the possibility that simpler solutions satisfying the condition of vanishing gradient may exist. In this work, we aim to identify such solutions, which we seek using numerical methods.

An additional advantage of this approach is its broader applicability. In scenarios where an explicit analytic expression for the potential is inaccessible—yet numerical evaluation remains feasible—our method retains its validity. Hence, the numerical strategy adopted here may prove effective even in cases where an exact analytic expression for the potential is out of reach.

\section{Numerical analysis}
	\subsection{Gradient descent and local analysis}
		%\paragraph{Gradient descent: implementation, sampling, time needed, loss}

@@

		\begin{figure}
			\centering
			\includegraphics[scale=0.35]{Figures/log_grad_V_hist.png}
			\quad
			\includegraphics[scale=0.35]{Figures/log_Vp4_hist.png}
			\caption{Left panel: Histogram of the logarithm of the norm of the gradient for each point. Right panel: Histogram of the distance of the value of the potential for each point.}
			\label{fig:Analyze_V_points}
		\end{figure}

@@

		\begin{figure}
			\centering
			\includegraphics[scale = 0.5]{Figures/Loss_grad_des.png}
			\caption{Evolution of the loss function during gradient descent and corresponding learning rate schedule, both plotted on a logarithmic scale. Dashed lines indicate epochs at which the optimizer was reinitialized.}
			\label{fig:loss_grad_des}
		\end{figure}

@@

		We can make a few comments on those plots. First, we observe non-trivial correlations in the data. Some of these may result from larger basins of attraction, such as in the $x_1/x_2$ plot. However, the structures seen in the $x_4/\tilde{x}_8$ or $x_4/x_{10}$ plots likely reflect genuine features of the manifold, and will show latter that this is indeed the case. We also note that all directions appear to be well populated. 
		%Furthermore, we can notice that a large majority of the points converge to the origin $\vec{0}$. This is not surprising, as this just mean that the round sphere solution has a larger basin of attraction, which sounds reasonable. We do not however attempt to shift the points along flat directions to depopulate the origin or repopulate other regions.

		%\paragraph{Local PCA: graphs of number of patches with given local dimension, and with varying size of local patch, discuss optimal patch size (we want at least few points in each direction)}

		Once the gradient descent has been completed and the flat direction sampled, the next step is to identify the structure of the underlying manifold. Our goal is to eventually obtain an analytical expression, not just a numerical description. Before applying symbolic regression to search for such an expression, we can first perform some exploratory analyses to better understand the data. Specifically, we aim to determine the dimensionality of the manifold and wether it consists of a single connected component or multiple disjoint components (e.g., two intersecting hyperplanes).

@@

		\begin{figure}[h!]
			\centering
			\includegraphics[scale=0.4]{Figures/Local_PCA.png}
			\caption{Results of the local PCA analysis. The $x$-axis shows the dimension inferred by the algorithm, and the $y$-axis indicates the proportion of points for which that dimension was found. Each curve corresponds to a different value of $k$ in the $k$-nearest neighbours.}
			\label{local_pca}
		\end{figure}

		We observe that for every choice of $k$, there is a prominent peak at $d = 3$, suggesting that the underlying manifold is three-dimensional. For $k = 5$, a noticeable fraction of points are assigned a dimensionality of 2. This can be attributed to the fact that if the true dimension is 3, then selecting only 5 neighbours may not sufficiently populate all three directions, leading the algorithm to underestimate the dimensionality for a fraction of the points.

@@

		%\paragraph{Clustering: HDBScan (only parameter: minimal number of points in a cluster), and graphs of 3d slices of the 5d space}

@@

		\begin{figure}[h!]
			\centering
			\hspace{-2cm}
			\includegraphics[scale=0.5]{Figures/3d_plot_x4_x8_x10.png}
			\includegraphics[scale=0.5]{Figures/3d_plot_x2_x4_x8.png}
			\caption{3D plots of the data in selected coordinates. Left: $(x_4, x_8, x_{10})$. Right: $(x_2, x_4, x_8)$.}
			\label{3dplots}
		\end{figure}

@@

	\subsection{Annealed Importance Sampling for polynomial symbolic regression}
	%\begin{itemize}[label=\textbullet]
		%\item We can convert the potential to a polynomial by converting the variables that appear in exponentials to logs.
	%\item As the potential is a polynomial, the solutions satisfy polynomials equations. Discuss that fact that we could get polynomials directly by taking the gradient of the potential, but those will be to complicated to express the vacuum in a usable way.  We search those using Annealed Importance Sampling.
